Proportional multi-state multiple-cohort life table model

### 2018-04-12 MSLT/MSLTdoc.Rmd
Added comparison graphs
Error when knit to pdf

```{r, echo=TRUE}
#Data preparation (generate age groups, change capitals to lower case)
##All upper cases to lower cases

@@

##Create data frames for Greater London and England population

@@

@@

```

@@

We define parameters for our model that are used multiple times. *p_age_cohort* defines the starting age for the cohorts. *p_sex* is for the cohort sex. *p_disease* specifies the modeled diseases. *p_intervention* is a hypothetical increase in MET minutes per week.

# ```{r, echo=TRUE}
# #5
# p_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
# p_sex <- c("males", "females")
# p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
# p_intervention_effect <- 100
# ```

####Data preparation

The following code applies to the data used as an example here. #6 transforms upper cases to lower cases in the sex variable. #7 and #8 is used to rename mortality_rate to mx (used in life table language). #9 is used to replace male and female in the exposure data to males and females to facilitate combination of data sets.

# ```{r, echo=TRUE}
# #6
# idata$sex <- tolower(idata$sex)
# #7
# idata$mx <- idata$mortality_rate
# #8
# idata$mortality_rate <- NULL
# #9
# edata$sex[edata$sex=="male"] <- "males"
# edata$sex[edata$sex=="female"] <- "females"
# ```

# ```{r, echo=TRUE}
# ##### Create new variable for 5_year population (age cohorts). Depends on the age-cohorts of interest (5-yrs here)
# 
# idata$five_year_population <- NA
# start_index <- 3
# index <- 1
# end_index <- 4
# for (i in 1:20){
#   if (i == 1)
#     index <- 1
#   else 
#     index <- ((i - 1)  * 5) + 1
#   if (index == 96){
#     end_index <- 5
#   }
#   cat("population sum ", start_index, "start index ", index, " and end index ", index + end_index, "\n")
#   idata$five_year_population[start_index] <- sum(idata$population[index:(index + end_index)])
#   start_index <- start_index + 5
# }
# 
# end_index <- 4
# start_index <- 104
# 
# for (i in 1:20){
#   if (i == 1)
#     index <- 102
#   else
#     index <- (((i - 1)  * 5) + 1) + 101
#   if (index == 197){
#     end_index <- 5
#   }
#   cat("population sum ", start_index, "start index ", index, " and end index ", index + end_index, "\n")
#   idata$five_year_population[start_index] <- sum(idata$population[index:(index + end_index)])
#   start_index <- start_index + 5
# }
# 
# ```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

>>>>

```{r, echo=TRUE}
#Data preparation (generate age groups, change capitals to lower case)

##All upper cases to lower cases

@@

##Create data frames for Greater London and England population to be later used as inputs for PMSLT calculations

@@

```{r, echo=FALSE}
###TO DO: loops for sex, cause and measure and interpolation of rates, care with interpolating rates, need to convert to log for interpolation using cubic spline. 
#FOR NOW THE CODE FOR GRAPH IS repetitive
####################MORTALITY RATES ALL CAUSE PER ONE

###############################Mortality rates###################################################################

@@

#############YLD RATES ALL CAUSE PER ONE

##################################yld rates all cause############################################################

##Select data from data frames for all cause yld rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause yld rates males 
totalyld_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Total YLDs rate")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Total YLDs rates difference males") +
  theme_classic()

print(totalyld_plot_compare_male)

##Select data from data frames for all cause yld females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause yld rates females 
totalyld_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Total YLDs rate")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Total YLDs rates difference females") +
  theme_classic()

print(totalyld_plot_compare_female)

################IHD
##############################Mortality rate ihd #############################################################

##Select data from data frames for ihd yld rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause yld rates males 
mx_ihd_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Mortality IHD")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "IHD mortality rates difference males") +
  theme_classic()

print(mx_ihd_plot_compare_male)

##Select data from data frames for ihd mortality rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause yld rates females 
mx_ihd_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Mortality IHD")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "IHD mortality rates difference females") +
  theme_classic()

print(mx_ihd_plot_compare_female)

##############################Prevalence rate ihd #############################################################

##Select data from data frames for ihd prevalence rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates males 
px_ihd_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Prevalence IHD")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "IHD prevalence rates difference males") +
  theme_classic()

print(px_ihd_plot_compare_male)

##Select data from data frames for ihd prevalence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates females 
px_ihd_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Prevalence IHD")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "IHD prevalence rates difference females") +
  theme_classic()

print(px_ihd_plot_compare_female)

##############################Incidence rate ihd #############################################################

##Select data from data frames for ihd prevalence rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates males 
ix_ihd_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Incidence IHD")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "IHD incidence rates difference males") +
  theme_classic()

print(ix_ihd_plot_compare_male)

##Select data from data frames for ihd prevalence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates females 
ix_ihd_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Incidence IHD")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "IHD incidence rates difference females") +
  theme_classic()

print(ix_ihd_plot_compare_female)

################ISCHEMIC STROKE
##############################Mortality rate ischemic stroke #############################################################

##Select data from data frames for ischemic stroke mortality rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot ischemic stroke mortality rates males

mx_is_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Mortality IHD")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Ischemic stroke mortality rates difference males") +
  theme_classic()

print(mx_is_plot_compare_male)

##Select data from data frames for ischemic stroke mortality rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot ischemic stroke mortality rates females 
mx_is_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Mortality ischemic stroke")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Ischemic stroke mortality rates difference females") +
  theme_classic()

print(mx_is_plot_compare_female)

##############################Prevalence rate ischemic stroke #############################################################

##Select data from data frames for ihd prevalence rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates males 
px_is_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Prevalence IHD")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Ischemic stroke prevalence rates difference males") +
  theme_classic()

print(px_is_plot_compare_male)

##Select data from data frames for ischemic stroke prevalence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates females 
px_is_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Prevalence ischemic stroke")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Ischemic stroke prevalence rates difference females") +
  theme_classic()

print(px_is_plot_compare_female)

##############################Incidence rate ischemic stroke #############################################################

##Select data from data frames for ishcemic stroke incidence rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates males 
ix_is_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Incidence ischemic stroke")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Ischemic stroke incidence rates difference males") +
  theme_classic()

print(ix_is_plot_compare_male)

##Select data from data frames for ischemic stroke incidence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot ischemic stroke incidence rates females 
ix_is_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Incidence ischemic stroke")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Ischemic stroke incidence rates difference females") +
  theme_classic()

print(ix_is_plot_compare_female)

################DIABETES MELLITUS
##############################Mortality rate diabetes mellitus #############################################################

##Select data from data frames for diabetes mellitus mortality rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot ischemic stroke mortality rates males

mx_DM_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Mortality diabetes mellitus")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Diabetes mellitus mortality rates difference males") +
  theme_classic()

print(mx_DM_plot_compare_male)

##Select data from data frames for diabetes mellitus mortality rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot diabetes mellitus mortality rates females 
mx_DM_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Mortality diabetes mellitus")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Diabetes mellitus mortality rates difference females") +
  theme_classic()

print(mx_DM_plot_compare_female)

##############################Prevalence rate diabetes mellitus #############################################################

##Select data from data frames for diabetes mellitus prevalence rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates males 
px_DM_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Prevalence diabetes mellitus")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Diabetes mellitus prevalence rates difference males") +
  theme_classic()

print(px_DM_plot_compare_male)

##Select data from data frames for diabetes mellitus prevalence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates females 
px_DM_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Prevalence diabetes mellitus")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Ischemic stroke prevalence rates difference females") +
  theme_classic()

print(px_DM_plot_compare_female)

##############################Incidence rate diabetes mellitus #############################################################

##Select data from data frames for diabetes mellitus incidence rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot incidence rate diabetes mellitus males 
ix_DM_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Diabetes mellitus")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Diabetes mellitus incidence rates difference males") +
  theme_classic()

print(ix_DM_plot_compare_male)

##Select data from data frames for diabetes mellitus incidence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot ischemic stroke incidence rates females 
ix_DM_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Incidence diabetes mellitus")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Diabetes mellitus incidence rates difference females") +
  theme_classic()

print(ix_DM_plot_compare_female)

#####Breast cancer
##############################Mortality rate breast cancer #############################################################

##Select data from data frames for breast cancer mortality rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "breast cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "breast cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot breast cancer mortality rates females 
mx_BC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Mortality breast cancer")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Breast cancer mortality rates difference females") +
  theme_classic()

print(mx_BC_plot_compare_female)

##############################Prevalence rate breast cancer #############################################################

##Select data from data frames for breast cancer prevalence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "breast cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "breast cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot all cause prevalence rates females 
px_BC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Prevalence breast cancer")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Breast cancer prevalence rates difference females") +
  theme_classic()

print(px_BC_plot_compare_female)

##############################Incidence rate breast cancer #############################################################

##Select data from data frames for breast cancer incidence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "breast cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "breast cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot breast cancer incidence rates females 
ix_BC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Incidence breast cancer")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Breast cancer incidence rates difference females") +
  theme_classic()

print(ix_BC_plot_compare_female)

################Colon and rectum cancer (DOSE RESPONSE FOR COLON CANCER?)
##############################Mortality rate colon and rectum cancer #############################################################

##Select data from data frames for colon and rectum cancer mortality rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot colon and rectum cancer mortality rates males

mx_CC_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Mortality colon and rectum cancer")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "colon and rectum cancer mortality rates difference males") +
  theme_classic()

print(mx_CC_plot_compare_male)

##Select data from data frames for colon and rectum cancer mortality rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot colon and rectum cancer mortality rates females 
mx_CC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Mortality colon and rectum cancer")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Colon and rectum cancers mortality rates difference females") +
  theme_classic()

print(mx_CC_plot_compare_female)

##############################Prevalence rate colon and rectum cancer #############################################################

##Select data from data frames for diabetes mellitus prevalence rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot colon and rectum cancer prevalence rates males 
px_CC_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Prevalence colon and rectum cancer")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "colon and rectum cancer prevalence rates difference males") +
  theme_classic()

print(px_CC_plot_compare_male)

##Select data from data frames for colon and rectum cancer prevalence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot colon and rectum cancer prevalence rates females 
px_CC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Prevalence colon and rectum cancer")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Colon and rectum cancer prevalence rates difference females") +
  theme_classic()

print(px_CC_plot_compare_female)

##############################Incidence rate colon and rectum cancer #############################################################

##Select data from data frames for colon and rectum cancer incidence rates males

ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot incidence rate colon and rectum cancer males 
ix_CC_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Colon and rectum cancer incidence rate")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "colon and rectum cancer incidence rates difference males") +
  theme_classic()

print(ix_CC_plot_compare_male)

##Select data from data frames for colon and rectum cancer incidence rates females

ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)

##Plot colon and rectum cancer incidence rates females 
ix_CC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
  geom_line(aes(color = "England"))+
  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
  labs(colour="Locations",x="Age",y="Incidence colon and rectum cancer")+
  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
  scale_color_manual(values = c("blue","red")) +
  labs (title = "Colon and rectum cancer incidence rates difference females") +
  theme_classic()

print(ix_CC_plot_compare_female)
```

plot(mx_plot_compare_male)
plot(mx_plot_compare_female)
plot(totalyld_plot_compare_male)
plot(totalyld_plot_compare_female)
plot(mx_ihd_plot_compare_male)
plot(mx_ihd_plot_compare_female)
plot(px_ihd_plot_compare_male)
plot(px_ihd_plot_compare_female)
plot(ix_ihd_plot_compare_male)
plot(ix_ihd_plot_compare_female)
plot(mx_is_plot_compare_male)
plot(mx_is_plot_compare_female)
plot(px_is_plot_compare_male)
plot(px_is_plot_compare_female)
plot(ix_is_plot_compare_male)
plot(ix_is_plot_compare_female)
plot(mx_DM_plot_compare_male)
plot(mx_DM_plot_compare_female)
plot(px_DM_plot_compare_male)
plot(px_DM_plot_compare_female)
plot(ix_DM_plot_compare_male)
plot(ix_DM_plot_compare_female)
plot(mx_CC_plot_compare_male)
plot(mx_CC_plot_compare_female)
plot(px_CC_plot_compare_male)
plot(px_CC_plot_compare_female)
plot(ix_CC_plot_compare_male)
plot(ix_CC_plot_compare_female)
plot(mx_BC_plot_compare_female)
plot(px_BC_plot_compare_female)
plot(ix_BC_plot_compare_female)
```

@@
